---
title: "Activity: Statistics of Gene Expression"
author: "Isha Thukral"
date: "Nov 9, 2021"
output: html_notebook
---

```{r}
# Load `tidyverse` and `DataComputing` packages
library(tidyverse)
library(mosaic)
library(dcData)
library(mosaicData)
library(esquisse)
```

```{r}
# Load the 'NCI60` data set into your RStudio environment from the `DataComputing` package

data("NCI60")
data("NCI60cells")
```

```{r}
Narrow <-
  NCI60 %>%
  tidyr::pivot_longer(cols = -Probe, names_to = "cellLine", values_to = "expression")

CellTypes <-
  NCI60cells %>%
  select(cellLine, tissue) %>%
  mutate(cellLine = gsub(pattern = "\\:", replacement = ".", x = as.character(cellLine)))

Narrow <- 
  Narrow %>%
  inner_join(CellTypes)
```

```{r}
Probe_TOP3A <-
  Narrow %>%
  filter(Probe == "TOP3A")
```

```{r}
SummaryStats <-
  Probe_TOP3A %>%
  group_by(tissue) %>%
  summarise(mn_expr = exp(mean(expression, na.rm = TRUE)))
```

```{r}
SummaryStats %>% 
  ggplot(aes(x = tissue, y = mn_expr)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```
```{r}
Probe_TOP3A %>% 
ggplot(aes(x = tissue, y = exp(expression))) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```
From Figure 18.1, the geom_point makes it difficult to see where the mean of each tissue lies. The points show the outliers, and a geom_bar would better compare the tissue in this data frame. If we implement a bar graph, we can exclude outliers from our judgment.
```{r}
    SummaryStats <-  
      SummaryStats %>%  
      mutate(tissue = reorder(tissue, mn_expr))
```

```{r}
SummaryStats <-
  Probe_TOP3A %>%
  group_by(tissue) %>%
  summarise(mn_expr = mean(expression,  na.rm = TRUE),
            se = sd(expression, na.rm = TRUE) / sqrt(n()))
```

```{r}
SummaryStats %>% 
  ggplot(aes(x = tissue, y = exp(mn_expr))) +
  geom_bar(stat = "identity", fill = "gray", color = NA) +
  geom_point(data = Probe_TOP3A, aes(x = tissue, y = exp(expression))) + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

```{r}
SummaryStats <- 
  SummaryStats %>%
  mutate(top = mn_expr + 2 * se,
         bottom = mn_expr - 2 * se) 
SummaryStats %>%
  ggplot(aes(x = tissue, y = exp(mn_expr))) +
  geom_bar(stat = "identity", alpha = 0.2) +
  geom_errorbar(aes(x = tissue, 
                    ymax = exp(top), 
                    ymin = exp(bottom)), width = 0.5) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```
```{r}
SummaryStats <- 
  SummaryStats %>%
  mutate(top = mn_expr + 2 * se,
         bottom = mn_expr - 2 * se) 
SummaryStats %>%
  ggplot(aes(x = tissue, y = exp(mn_expr))) +
  geom_point(data = Probe_TOP3A, aes(x = tissue, y = exp(expression))) +
  geom_errorbar(aes(x = tissue, 
                    ymax = exp(top), 
                    ymin = exp(bottom)), width = 0.5) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

```{r}
# customize a user-defined function called `r2`
r2 <- function(data) {
  mosaic::rsquared(lm(data$expression ~ data$tissue))
}
```

```{r}
ProbeR2 <- 
  Narrow %>% 
  group_by(Probe) %>%
  dplyr::do(probe_rsq = r2(.)) %>%
  mutate(r2 = unlist(probe_rsq))
```

```{r}
Actual <-
  ProbeR2 %>%
  arrange(desc(r2)) %>%
  head(30) 
```

```{r}
Actual %>%
  ggplot(aes(x = reorder(Probe, desc(r2)), y = r2)) +
  geom_point() + 
  xlab("Probe") + 
  ylab("R-squared") + 
  ggtitle("Probes with large R-squared for expression level explained by tissue type") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
New_Probe<-
  Narrow %>%
  filter(Probe=="MYB")

SummaryStats <-
  New_Probe %>%
  group_by(tissue) %>%
  summarise(mn_expr = mean(expression,  na.rm = TRUE),
            se = sd(expression, na.rm = TRUE) / sqrt(n())) %>%
  mutate(top = mn_expr + 2 * se,
         bottom = mn_expr - 2 * se) 

SummaryStats %>%
  ggplot(aes(x = tissue, y = exp(mn_expr))) +
  geom_point(data = New_Probe, aes(x=tissue, y=exp(expression))) +
  geom_errorbar(aes(x = tissue, 
                    ymax = exp(top), 
                    ymin = exp(bottom)), width=0.5) +
  theme(axis.text.x = element_text(angle = 30, hjust=1)) 
```
After testing out a couple of the probes I have observed that probes with high R^2 probe have one point that doesn't show 0 and makes the R^2 larger. Wherease points with lower R^2 have more points that aren't 0 such as MYB.

```{r}
NullR2 <-
  Narrow %>%
  group_by(Probe) %>%
  mutate(expression = mosaic::shuffle(expression)) %>%
  group_by(Probe) %>%
  do(r2 = r2(.)) %>%
  mutate(r2 = unlist(r2))
```
```{r}
ProbeR2 %>%
  ggplot(aes(x = r2)) + 
  geom_density(fill = "gray30", color = NA) +
  geom_density(data = NullR2, aes(x = r2), 
               fill = "gray80", alpha = .75, color = NA)
```
```{r}
Null <-
  NullR2 %>% 
  arrange(desc(r2)) %>%
  head(30) 

# append the 30 highest `Null` values to the `Actual` data
Actual$null <- Null$r2

Actual %>%
  ggplot(aes(x = reorder(Probe, desc(r2)), y = r2)) +
  geom_point() + 
  geom_point(aes(y = null), color = "gray50") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


