---
title: "Final Project: Initial EDA"
output: html_notebook
---

```{r}
# clean up the RStudio environment 
rm(list = ls())

# load all packages here: `mosaic`, `tidyverse`, `lubridate`, and all others used
library(tidyverse)
library(mosaicData)
library(esquisse)
library(dcData)
library(lubridate)
library(ggplot2)
library(readr)
```

```{r}
 # Tip: Scroll right if browser settings cut off the URL
cities_internet_prices_historical_24_10_2021 <- read_csv("cities_internet_prices_historical.24-10-2021.csv")
    
 # Tip: Scroll right if browser settings cut off the URL
GDP <- read_csv("GDP.csv")
```

### Data Intake

```{r}
summary(cities_internet_prices_historical_24_10_2021)
```

```{r}
glimpse(cities_internet_prices_historical_24_10_2021)
```

```{r}
head(cities_internet_prices_historical_24_10_2021)
```


```{r}
tail(cities_internet_prices_historical_24_10_2021)
```
```{r}
summary(GDP)
```


```{r}
glimpse(GDP)
```


```{r}
head(GDP)
```
```{r}
tail(GDP)
```

### Research Question (Need to EDIT!!!!!!!!!!!!!!!!!!!!!!)

We are going to use two datasets: internet price and GDP. We are going to investigate the relationship between internet price and GDP by joining two datasets. In hypothesis, we expect that countries with higher GDP have higher price of internet.

### Where the data from

**WHERE:** We found our two datasets from [kaggle](https://www.kaggle.com/datasets).

[World cities average inter prices. 2010-2020](https://www.kaggle.com/cityapiio/world-cities-average-internet-prices-2010-2020).

[GDP per capita all countries](https://www.kaggle.com/nitishabharathi/gdp-per-capita-all-countries).

**HOW:** Since all of our group members had interest in joining datasets and in economics, we decided to look for GDP data with country and year variable and look for any data that has country and year variables in it.


### Examine the data source (Need to EDIT!!!!!!!!!!!!!!!!!!!!!!)

Report what you know (or don't know) about who/what/when/where/why/how the data were originally collected (note: this is NOT about how you found it, etc)


choose a few (3-4) key variables to describe (variable types, coding, missingness, summary statistics/plots)

Heejun's Suggestion: We are going to use mainly 4 variables: GDP per countries, internet price per countries, year, and country names. Since internet price is not for country but for cities, we are going to use coding to find average internet price for each countries. Variable types are basically numbers, but there are some "null" values that we are going to remove by coding. For summary statistics/plots, we are going to calculate R squared score to see the strength of the relationship between two datasets and are going to visualize a plot including data points, regression line, or r^2 value by using ggplot() or equisse().

### Discover/Explore features that may influence modeling decisions (both sources)

Need to fill

### Address research question based on observations from your data sources

Need to fill

