---
title: "STAT 184 Homework Exercises" 
subtitle: "DataComputing Chapters 10"
author: "Isha Thukral"
output: html_notebook
---

## Assigned Exercises

- DC Chapter 10 Exercises: 10.1, 10.2, 10.3, 10.4, 10.5, 10.6, 10.7 

## Front Matter

```{r}
# clean up workspace environment
rm(list = ls())

# all packages used for the assignment
library(mosaic)
library(tidyverse)
library(mosaicData)
library(esquisse)
library(dcData)

```



## Chapter 10 Exercises

### Problem 10.1

1. str(): summary function
2. group_by():Data verbs
3. rank(): transformation  
4. mean():reduction   
5. filter(): Data Verb 
6. summary():summary function 
7. summarise(): data verb 
8. glimpse():summary function  

### Problem 10.2
1. mutate()
2. arrange(desc())
3. filter() specified for criteria 
4. select() with column A and X specified 
5. group_by(), summarise(max())

### Problem 10.3
1. summarise(count=n())
2. names()
3. help("codebook")
4. library(x)
5. group_by() 

### Problem 10.4
1. Data verb: arrange(count)
2. Data verb: filter(sex != "M")
3. Data verb: filter(sex != "F")
4. Data verb: summarise(total = sum(count))
5. Data verb: select(name, count) 

### Problem 10.5

1. First filter() rows to find values that have first and second choice the same. Than use the n() to display the total number of cases.
```{r}
Minneapolis2013%>%
  filter(First == Second)%>%
  summarise(count = n())
```
2. Similiar to question one we would use the implement function to find cases where first and second are the same. Then group_by (for columns) to get the number of rows for a specific name. Next using n() shows case numbers for first and second. Finally arrange in descending order and retrieve top 3 by using head. 
```{r}
Minneapolis2013%>%
  filter(First == Second)%>%
  group_by(First, Second)%>%
  summarise(choices = n())%>%
  arrange(desc(choices))%>%
  head(3) 
```
3. Using filter function to gather values with "OLE SAVIOR" case. Then group_by second choice column to get total of each name and display n(). Finally put in descending order and use head to get top 3. 
```{r}
Minneapolis2013%>%
  filter(First == "OLE SAVIOR")%>%
  group_by(Second)%>%
  summarise(choices = n())%>%
  arrange(desc(choices))%>%
  head(3)
```

### Problem 10.6
1. 
```{r}
diamonds%>%
  group_by(color)%>%
  summarise(average = sum(mean(carat)))%>%
  arrange(desc(average))
```
2. 
```{r}
diamonds%>%
  group_by(clarity)%>%
  summarise(average = sum(mean(table/carat)))%>%
  arrange(desc(average)) 
```

### Problem 10.7
```{r}
Minneapolis2013%>%
  group_by(First)%>%
  summarise(count = n())%>%
  mutate(First = ifelse(count < 5000, "minor", First))
  
```


