---
title: "STAT 184 Homework Exercises" 
subtitle: "DataComputing Chapters 15"
author: "Isha Thukral"
output: html_notebook
---

## Assigned Exercises
 
- DC Chapter 15 Exercises: 15.1, 15.2, 15.3, 15.4, 15.5

## Front Matter

```{r}
# clean up workspace environment
rm(list = ls())

# all packages used for the assignment
library(mosaic)
library(tidyverse)
library(mosaicData)
library(esquisse)
library(dcData)
library(lubridate)
```

## Chapter 15 Exercises

### Problem 15.1
```{r}
mosaicData::CPS85 %>%
  ggplot(aes(x=sex,y=wage)) +
  geom_boxplot(aes(fill=sex))
```


### Problem 15.2
1.65 meters

2.female = 1.62 meters, male = 1.77 meters

3.BMI for 40 y/o = 26, BMI for 40 y/o = 28

### Problem 15.3
```{r}
mosaicData::CPS85 %>%
  ggplot(aes(x=educ, y=wage, color = sex)) +
  geom_point() + stat_smooth(method = "lm") +
  ylim(0,15)
```
### Problem 15.4
```{r}
mosaicData::Galton %>%
  ggplot(aes(x=height,y=mother)) +
  stat_density2d() +
  facet_grid(~sex)+
  theme_bw()
```
### Problem 15.5
1. The following data doesn't show the ZIP codes with high elderly population tend to have high child populations, as the scattered data where the increase in fraction of 5 year olds and increase in fraction of 65 and older can be seen simultaneously.

2. Looking at the data it seems to be consistent as the confidence bands have a broad variety of data points and possible a true relationship. Plus as it could not be a no-relationship. The line are pretty much parallel to x-axis and if more points gathered can lead to non-level line.

### You decide to get more data: study 80 ZIP codes. Here’s the result.

1. The horizontal line is not consistent with the data, as its in downward trend.

2. For 80 zipcodes to 20 zipcodes the ratio would be 1:2.    

3. Yes it is consistent and seems right because (1/√20) * (1/√80) equals 2. And the answer is same for the same from the previous question. 

